### 2015-05-14 whoop.tex
some space hack

\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver (see \S\ref{evaluation}), we quickly exhausted the memory limits of our experimental platform as the driver exhibited recursion, which cannot be fully inlined and blows up.

To make our analysis scalable without sacrificing precision, and to cater for recursion, we use the Houdini~\cite{flanagan2001houdini} invariant inference algorithm to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants.  Given a pool of candidate invariants, Houdini repeatedly tries to verify each procedure with respect to its current candidate summary.  If verification fails due to an incorrect candidate, this candidate is discarded.  The process repeats until a fixpoint is reached, whereby the largest conjunction of the candidate invariants that can be proven to hold reamins.  The remaining candidates, which are indeed invariants, can then be assumed during further analysis of the program.

The Houdini algorithm does not generate the initial pool of candidate invariants; these are given to the algorithm as a starting point.  \whoop generates an initial pool of candidates using a set of heuristics.  The idea is to automatically generate likely program invariants based on syntactic patterns extracted from an inter-procedureal pass over the code for an entry point.  We give two examples; for clarity we use notation from the simple shared variable concurrent programming model of \S\ref{sec:symbolicpairwise}.  If we observe syntactically that procedure $f$ of entry point $T$ may write to, but does not read from, shared variable $s$.  Then when instrumenting $T$ with symbol $i$, we guess $s \in \hasbeenwritten{i}$ and $s \notin\hasbeenread{i}$ as postconditions for $\instrument{f}{i}$.  These guesses may be incorrect, for instance if the potential write to $s$ turns out to be in dead code, or if a read from $s$ has already been issued on entry to $\instrument{f}{i}$.  Similarly, if syntactic analysis indicates that $f$ may unlock mutex $m$, we guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$; this guess may be wrong, e.g.\ if the unlock operation is not reachable or if a subsequent lock operation acquires the mutex again.  We stress that guessing of incorrect candidate invariants does not compromise the soundness of verification: \whoop is free to speculatively generate candidates that are later deemed to be incorrect, and thus discarded, by Houdini.  The balance we try to strike is to have \whoop generate sufficient candidates to enable precise lockset analysis, without generating so many candidates that the speed of the Houdini algorithm is prohibitively slow.

In practice, the candidates are generated at the Boogie level, with respect to SMACK-generated memory regions and referencing the auxiliary variables introduced by our watchdog race instrumentation.

>>>>

\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop, we experimented with analyzing recursion-free drivers using full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver (see \S\ref{evaluation}), we quickly exhausted the memory limits of our experimental platform as the driver exhibited recursion, which cannot be fully inlined.

To make our analysis scalable without sacrificing precision, and to cater for recursion, we use the Houdini~\cite{flanagan2001houdini} invariant inference algorithm to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants.  Given a pool of candidates, Houdini repeatedly attempts to verify each procedure with respect to its current candidate summary.  If verification fails due to an incorrect candidate, this candidate is discarded.  The process repeats until a fixpoint is reached, whereby the largest conjunction of the candidate invariants that can be proven to hold remains.  The remaining candidates, which are indeed invariants, can then be assumed during further analysis of the program.

Houdini does not generate the initial pool of candidates: \whoop generates them using a set of heuristics, and passes them to Houdini as a starting point.  The idea is to automatically generate likely program invariants based on syntactic patterns extracted from an inter-procedural pass over the code for an entry point.  We give two examples; for clarity we use notation from the simple shared variable concurrent programming model of \S\ref{sec:symbolicpairwise}.  If we observe syntactically that procedure $f$ of entry point $T$ may write to, but does not read from, shared variable $s$, then when instrumenting $T$ with symbol $i$, we guess $s \in \hasbeenwritten{i}$ and $s \notin\hasbeenread{i}$ as postconditions for $\instrument{f}{i}$.  These guesses may be incorrect, for instance if the potential write to $s$ turns out to be in dead code, or if a read from $s$ has already been issued on entry to $\instrument{f}{i}$.  Similarly, if syntactic analysis indicates that $f$ may unlock mutex $m$, we guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$; this guess may be wrong, e.g.\ if the unlock operation is not reachable or if a subsequent lock operation acquires the mutex again.  We stress that guessing incorrect candidate invariants does not compromise the soundness of verification: \whoop is free to speculatively generate candidates that are later deemed to be incorrect, and thus discarded, by Houdini.  The balance we try to strike is to have \whoop generate sufficient candidates to enable precise lockset analysis, without generating so many candidates that the speed of the Houdini algorithm is prohibitively slow.

In practice, the candidates are generated at the Boogie IVL level, with respect to SMACK-generated memory regions and referencing the auxiliary variables introduced by our watchdog race-checking instrumentation.
